---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# carbcontpred

<!-- badges: start -->
<!-- badges: end -->

The first goal of ````carbcontpred```` is to build and cross-validate a linear regression model for a data set containing the spectral reflectance and organic carbon content of soil samples, and then use it to predict the organic carbon content based on reflectance. The second goal of ````carbcontpred```` is to use the predictions from the linear regression model to predict the organic carbon content of a Region of Interest (ROI) data table containing the spectral reflectance of spatial points on a map.

## Installation

You can install the development version of ````carbcontpred```` from [GitHub](https://github.com/KangyuSo/carbcontpred) with:

``` r
install.packages("remotes", repos = "http://cran.us.r-project.org")
install.packages("knitr", repos = "http://cran.us.r-project.org")
install.packages("rmarkdown", repos = "http://cran.us.r-project.org")
library(remotes)
library(knitr)
library(rmarkdown)

remotes::install_github("KangyuSo/carbcontpred", 
                         build_vignettes = TRUE)
library(carbcontpred)
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## Install and load the packages required to read the "carbcontpred" package
install.packages("remotes", repos = "http://cran.us.r-project.org")
install.packages("knitr", repos = "http://cran.us.r-project.org")
install.packages("rmarkdown", repos = "http://cran.us.r-project.org")
library(remotes)
library(knitr)
library(rmarkdown)

## Install and load the "carbcontpred" package
remotes::install_github("KangyuSo/carbcontpred", 
                         build_vignettes = TRUE)
library(carbcontpred)

## Load the data set 
data <- data.frame(ref_dry_avg)

## Create a linear regression model with "Reflectance" as the predictor variable, "Organic_Carbon" as the response variable, a test size of 70%, 10 folds for the cross-validation, and "Predicted_Organic_Carbon" as the output column
pred_c <- cpred(data, "Reflectance", "Organic_Carbon", 0.7, 10, "Predicted_Organic_Carbon")
 
## View the results
View(pred_c)
   
## Load the ROI data file "ROI_Predicted_Carbon.csv" and apply min-max normalization on the spectral reflectance values 
## Extract an equation for predicting organic carbon content using reflectance based on the results of the "cpred" function and use it to predict the organic carbon content of the "ROI_Predicted_Carbon.csv" file using its reflectance values
## Update the ROI data file with the predictions and save it as a CSV file output titled "ROI_Predicted_Carbon"
mappred_c <- mappred(pred_c, "Reflectance", "Predicted_Organic_Carbon", system.file("extdata", "Reflectance_ROI.csv", package = "carbcontpred"), "ROI_Predicted_Carbon", "ROI_folder", "carbcontpred")

## View the first 10 rows of the resulting CSV file
head(read.csv(system.file("ROI_folder", "ROI_Predicted_Carbon.csv", package = "carbcontpred")), 10)
```