---
title: "Process pred_c File"
author: "Kangyu So"
date: "2023-12-17"
output: html_document
---

I have an `.RData` file titled `red_dry_avg` which was the output of `ref_dry_avg.R`. The file contains data on post-dried soil samples regarding their sample ID and their average spectral reflectance between the wavelengths 350 nm and 2500 nm. The soil samples were collected from the Hudson Bay Lowlands.

To increase the shareability of the data I will first read the data set, then convert it to an `.RData` file, and then document it. This will enable the data set to be bundled with package `carbcontpred`.

Install and load the package needed to preprocess the data:
```{r}
remotes::install_github("KangyuSo/carbcontpred", 
                        build_vignettes = TRUE)
library(carbcontpred)
```

Read the `red_dry_avg`` file as a data frame:
```{r}
data <- data.frame(ref_dry_avg)
```

The data frame consists of 432 observations of 3 variables. Get a summary of the data frame:
```{r}
summary(data)
```

We want to get predict the organic carbon content of the soil samples using their spectral reflectance and measured organic carbon content. We will create a linear regression model where the reflectance is the predictor variable and the organic carbon content is the response variable. The model will also have a test size of 70% and 10 folds for the cross-validation. The model will make predictions for the organic carbon content of the soil samples and output them with the other variables from the original data frame under a new column with the header `Predicted_Organic_Carbon`:
```{r}
pred_c <- cpred(data, "Reflectance", "Organic_Carbon", 0.7, 10, "Predicted_Organic_Carbon")
```

Load the summary of the new data frame and ensure that there is no errors in the data:
```{r}
summary(pred_c)
```

Save to data folder:
```{r}
usethis::use_data(pred_c, overwrite = TRUE)
```
