---
title: "ref_dry_avg"
author: "Kangyu So"
date: "2023-11-06"
output: html_document
---

I have two CSV files titled "SoilReflectanceDry_v01" and "SoilMetrics_v01". The "SoilReflectanceDry_v01" file contains data on the spectral reflectance of post-dried soil samples between the wavelengths 350 nm and 2500 nm in increments of 1 nm. The soil samples were collected from the Hudson Bay Lowlands. The "SoilMetrics_v01" file contains biophysical and biochemical metrics of the soil samples. This includes the organic carbon content of the soil samples by percent dry weight.

To increase the shareability of the data I will first read the two data sets into one. I will then convert the data set to an `.RData` file and then document it. This will enable the data set to be bundled with package `carbcontpred`.

Install and load the package needed to preprocess the data:
```{r}
install.packages("prospectr")
library(prospectr)
```

Read the "SoilReflectanceDry_v01" file as a data frame:
```{r}
data_ref <- read.csv("data-raw/SoilReflectanceDry_v01.csv")
```

The data frame consists of 2151 observations of 433 variables. Get a summary of the data frame:
```{r}
summary(data_ref)
```

The data was collected by Kangyu So, Dr. Alemu Gonsamo, and Dr. Cheryl A. Rogers from McMaster University. The data was collected from 18 sites near the Cree First Nations communities in the Hudson Bay Lowlands region of northern Ontario, Canada. The data was collected during the month of July 2023 and spectral measurements were collected during the month of August 2023. 

The reflectance spectra is not normalized so I will calculate continuum removal for the data frame:
```{r}
ref_dry <- as.data.frame(apply(data_ref[,-1], MARGIN=2, FUN =continuumRemoval, wav = data_ref$Wavelength))
ref_dry <- cbind(Wavelength = data_ref$Wavelength, ref_dry)
```

At this stage, the data is arranged so that its definitions are composed of the wavelength and the ID of all of the soil samples, totaling 433 definitions. There are too many data definitions right now so the data set must be changed into a more acceptable format. For this example, we will look at the entire wavelength range of 350 nm to 2500 nm for all of the samples.

Define the start_row to be the lower end of the wavelength range you are interested in and end_row to be the higher end, which would be 350 and 2500 respectively:
```{r}
start_row <- 350
end_row <- 2500
```

Select the rows to be the wavelengths you are interested in, which would be 350 to 2500 respectively. Select the columns for your subset to be the samples you are interested in, which would be all of the samples.
```{r}
ref_dry <- ref_dry[start_row:end_row, 2:ncol(ref_dry)]
```

Calculate the average reflectance for each column in the data set excluding NA values:
```{r}
avg_values <- colMeans(ref_dry, na.rm = TRUE)
```

Create a new data set that contains a column with the ID of the soil samples and another column with the average reflectance values, which will be named "Sample_ID" and "Reflectance" respectively: 
```{r}
ref_dry_avg <- data.frame(
  "Sample_ID" = colnames(ref_dry),
  "Reflectance" = avg_values
)
```

Read the "SoilMetrics_v01" file as a data frame:

```{r}
data_met <- read.csv("avgref/data-raw/SoilMetrics_v01.csv")
```

The data frame consists of XXX observations of XXX variables. Get a summary of the data frame:
```{r}
summary(data_met)
```

The data was collected by Kangyu So, Dr. Alemu Gonsamo, and Dr. Cheryl A. Rogers from McMaster University. The data was collected from 18 sites near the Cree First Nations communities in the Hudson Bay Lowlands region of northern Ontario, Canada. The data was collected during the month of July 2023 and biophysical measurements were collected during the month of August 2023. Biochemical measurements were done by the University of Guelph's Agriculture and Food Laboratory.

Install and load the package needed to preprocess the data:
```{r}
install.packages("dplyr")
library(dplyr)
```

Merge the organic carbon data from the "data_met" dataset to the "ref_dry_avg" dataset based on the ID of the soil samples
```{r}
ref_dry_avg <- ref_dry_avg %>%
  left_join(data_met %>% select(Sample_ID, OC), by = "Sample_ID") %>%
  rename(Organic_Carbon = OC)
```

Load the summary of the new data frame and ensure that there is no missing data:
```{r}
summary(ref_dry_avg)
```

Save to data folder:
```{r}
usethis::use_data(ref_dry_avg, overwrite = TRUE)
```