---
title: "carbcontpred-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{carbcontpred-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# carbcontpred: Predicting Organic Carbon Content Soil Samples Based on Their Spectral Reflectance in Hudson Bay Lowlands, Ontario, Canada

This package contains objects that are sourced from a 2022 McMaster Remote Sensing Laboratory research project and objects created to facilitate the use and analysis of the fieldwork data. The project recorded a wide range of biochemical, biophysical, and spectral metrics of 432 soil samples that were collected from the Hudson Bay Lowlands (N 51.220318, W 80.556189) and measured and analyzed at McMaster University and the University of Guelph. A portion of the project’s data has been cleaned and packaged for easy use in an ````R```` environment.

The project was funded by the Natural Sciences and Engineering Research Council of Canada (NSERC) through the Discovery Grants program (RGPIN-2020-05,708). The corresponding author is Kangyu So^[McMaster University, sok11@mcmaster.ca]. The full list of authors can be found here^[Authors: Kangyu So (McMaster University, sok11@mcmaster.ca), Cheryl A. Rogers (McMaster University, rogerc11@mcmaster.ca), Alemu Gonsamo (McMaster University, gonsamoa@mcmaster.ca), Yiyao Li (McMaster University, liy571@mcmaster.ca), James Snider (World Wildlife Fund Canada, jsnider@wwfcanada.org), Clare Wark (World Wildlife Fund Canada, cwark@wwfcanada.org)].

# What Is This Package?

The package ````carbcontpred```` contains objects that are sourced from a 2022 McMaster Remote Sensing Laboratory research project and objects created to facilitate the use and analysis of the fieldwork data. This package includes data on the spectral reflectance and organic carbon content of soil samples after they have been dried in an oven for 48 hours. The package also includes the function ````cpred```` for building a linear regression model to make organic carbon content predictions, and an ````R```` Markdown file on pre-processing the data for ````cpred````.

The raw data from the research project will eventually be freely available to the public through a repository on GitHub, but it is bulky and organized in a complicated manner. ````carbcontpred```` pre-processes the data in ````R````, extracting the portion of the data that is useful for developing predictive models for carbon stock in soil samples. The package also includes functions that can be used to build predictive models or be tuned in the future by the authors or an open-source community.

# Set-Up

## Installation

```{r eval = FALSE} 
install.packages('carbcontpred', repos = "https://github.com/KangyuSo/my-GEO712-repository/tree/main/carbcontpred")
```

## Pre-Processing Data

The 2022 McMaster Remote Sensing Laboratory research project collected soil samples from the Hudson Bay Lowlands region in northern Ontario, Canada (coordinates for the place!!!!). Included within are two csv files containing data for the soil samples. The first csv file contains data on the spectral reflectance of post-dried soil samples between the wavelengths 350 nm and 2500 nm in increments of 1 nm. 

The reflectance spectra are not normalized, so continuum removal must be applied. Which is done using the following chunk:

```{r eval = FALSE} 
ref_dry <- as.data.frame(apply(data_ref[,-1], MARGIN=2, FUN =continuumRemoval, wav = data_ref$Wavelength))
ref_dry <- cbind(Wavelength = data_ref$Wavelength, ref_dry)
```

The table is still arranged so that its definitions are composed of the wavelength and the ID of all of the soil samples, totaling 433 definitions. This is too many data definitions right now and we are only interested in the average reflectance of a given wavelength range. 

We will first calculate the average using the following chunk:

```{r eval = FALSE} 
start_row <- 350
end_row <- 2500
ref_dry <- ref_dry[start_row:end_row, 2:ncol(ref_dry)]
avg_values <- colMeans(ref_dry, na.rm = TRUE)
```

Next, we will create a new dataset that records the sample ID and average reflectance in a more acceptable format:

```{r eval = FALSE} 
ref_dry_avg <- data.frame(
  "Sample_ID" = colnames(ref_dry),
  "Reflectance" = avg_values
)
```

Now, we have a dataset with our sample ID and average reflectance, the latter of which will be the independent variable for the linear regression model. We need to add our dependent variable to this dataset, which would be the organic carbon content. This data is stored in ````SoilMetrics_v01.csv```` in the ````\data-raw folder````.

We will merge the organic carbon content data from this dataset to the dataset containing the sample ID and average reflectance using the following chunk:

```{r eval = FALSE} 
ref_dry_avg <- ref_dry_avg %>%
  left_join(data_met %>% select(Sample_ID, OC), by = "Sample_ID") %>%
  rename(Organic_Carbon = OC)
```

Finally, our data is fully processed and ready to build linear regression models using the ````cpred```` function.

See the process-ref_dry.Rmd file for complete details on how the raw data was processed for use in the ````carbcontpred```` package.

# Function Overview:

The ````cpred```` function builds a linear regression model using spectral reflectance as the independent variable and organic carbon content as the dependent variable and uses it to predict the organic carbon content of soil samples based on their reflectance. The function will utilize a dataset containing the spectral reflectance and organic carbon content of soil samples. In this package, that dataset will be ````ref_dry_avg.rda````.

The cpred function can be applied using the following chunk:

```{r eval = FALSE} 
data <- data.frame(ref_dry_avg)
cpred <- function(data, reflectance, carbon, test_size, folds, prediction)
```

For the chunk above, the parameter ````data```` represents the ref_dry_avg dataset. The parameter ````reflectance```` represents the average at a given wavelength range. The parameter ````carbon```` represents organic carbon content by percent dry weight. The parameters ````test_size```` and ````folds```` represent the test size for the model and the number of folds for the cross-validation respectively. The parameter ````prediction```` represents the name for the predicted organic carbon content output.

The return of the function is a new column added to the previous dataset which contains the predicted organic carbon content values using a cross-validated linear regression model.

See the ````README.Rmd```` files for additional details on how to utilize ````cpred````. See the ````cpred.R```` file for additional information about the function.

# Contributing to the carbcontpred Package

The purpose of this ````R```` package is to make the soil spectral reflectance and organic carbon content data from a 2022 McMaster University Remote Sensing Laboratory project easily available for predictive modeling in an ````R```` environment. Currently, the package only uses the reflectance and organic carbon content data of the post-dried soil samples, but we invite others from the community to request additional metrics from the original project dataset, report issues such as spelling errors, and contribute to the ````carbcontpred```` package. To contribute to the package, create a branch in the repository and submit a pull request for your contribution. Contributions will be reviewed before approval and full credit will be given.



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(carbcontpred)
```
